---
title: "modeltime"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Library

```{r LoadPackage,warning=FALSE,message=FALSE}
library(tidyverse)
library(ggplot2)
library(stringr)
library(lubridate)
library(ggthemes)
library(tidyquant)
library(forecast)
library(modeltime)
library(gridExtra)
library(urca)
library(tidymodels)
```


# Loading Epias Energy Consumption Data

```{r}
data = read.csv("D:/2022_yedek/Tolga/data/elektrik_tuketim.csv",check.names = F)
rbind(head(data,5),tail(data,5))
```


* The data shows hourly electricity consumption from 2015 to 2022.


* First let's prepare the historical column in the data according to the format we can analyze

```{r}
data$Tarih = as.Date(data$Tarih,format = "%d.%m.%Y") 
data$HourlyDate = seq(as.POSIXct("2015-12-31 00:00:00")
                      ,as.POSIXct("2022-05-14 16:00:00")
                      ,by = "hour")   # Create new date column with hour and date

colnames(data)[1:3] = c("Tarih","Saat","Tuketim")
data$Tuketim = sub(".", "",
                   data$Tuketim,
                   fixed = TRUE) # change some character problem from numeric data
data$Tuketim = sub(",",".",
                   data$Tuketim,
                   fixed = TRUE) # another one..

data$Tuketim = as.numeric(data$Tuketim) # finally transforming numerical data


ggplot(data,aes(x = HourlyDate,y = Tuketim))+
  geom_line() # first look consumption data
```
* One date has zero consumption value. Let's see what date it is.


```{r}
data%>%filter(Tuketim == 0)
```

* We have a zero value on 2016-03-27 02:00. I did some research but I don't know why.

* Anyway, now let's look at what hours the consumption increases.

```{r,fig.height=9,fig.width=15}
ggplot(data,aes(x = HourlyDate,y = Tuketim,colour = Saat))+
  geom_line()+
  scale_fill_viridis_b()+
  theme_economist()# line plot with hour
```

*The blue line shows midday hours. 
*The hours with the highest consumption are midday hours.
*This may be due to the fact that it is both the highest temperature and working hours.
*And the green line shows the hours between 04:00 and 10:00.
These are the hours with the lowest consumption. The reason for this may be reasons such as the inactivity of people.


```{r,warning=FALSE,message=FALSE,fig.height=8,fig.width=18}
data%>%group_by(Saat)%>%summarise(OrtalamaTuketim = mean(Tuketim))%>%
  ggplot(aes(x = Saat,y = OrtalamaTuketim,fill = Saat))+geom_bar(stat = "identity")+
  theme_fivethirtyeight()+
  ggtitle(label = "Mean Hourly Consumption")


data = data%>%mutate(Day = day(Tarih),
              Month = month(Tarih),
              Year = year(Tarih))
g = data%>%group_by(Year,Month,Day)%>%
  summarise(Mean = mean(Tuketim)) # Create mean daily consumption data.

g$Tarih = as.Date(paste0(g$Year,"-",g$Month,"-",g$Day),format = "%Y-%m-%d")
```

* In the average hourly consumption graph, there is a decrease after 00:00 until 06:00. After 06:00, the amount of consumption increases continuously and the maximum point is 11:00.

## Creating daily mean consumption data

```{r}
ggplot(g,aes(x = Tarih,y = Mean,color = "red"))+
  geom_line()+
  theme_economist()+
  ggtitle(label  = "Mean Daily Electric Consumption from EPİAS")+
  theme(legend.position = "none")
```

And now we look sma and coloring by month.

```{r,fig.height=8,fig.width=15}
ggplot(g,aes(x = Tarih,y = Mean,colour = Month))+
  geom_line()+
  scale_fill_viridis_b()+
  geom_ma(ma_fun = EMA,n = 48,size = 1.5,color = "red")+
  theme_economist_white(horizontal = T)

```
Consumption always peaks as we approach the last months of the year.
This may be caused by the cooling consumption that occurs with the arrival of the summer season and the increase in temperatures.
However, in some years (for example, 2020), the months when consumption is at its peak coincide with the winter season.

```{r,fig.height=8,fig.width=15}
g%>%group_by(Month)%>%summarise(OrtalamaTuketim = mean(Mean))%>%
  ggplot(aes(x = month(Month,label = T),y = OrtalamaTuketim,fill = Month))+geom_bar(stat = "identity")+
  ggtitle(label = "Mean Monthly Consumption")+
  theme_fivethirtyeight()
```
The months with the highest consumption are July and August, when the summer temperature is also the highest. 
In addition, the lowest month belongs to the month of May, which we can call the transition month between winter and summer.

# Do forecast!

```{r,fig.height=15,fig.width=20,warning=FALSE,message=FALSE}
ts_data = ts(g$Mean,
             start = c(2015,yday(g$Tarih[1])),
             frequency = 365) # Creating ts object for forecast package.




plot_with = function(ts_object,seasonality) # Create function and show acf,pacf and data plot!
{
  if(seasonality=="True")
    {
      s1 = ggseasonplot(ts_data,size = 1)+theme_solarized()
    }
 
  

  p1 = autoplot(ts_object,color = "red")+
    autolayer(SMA(ts_object,n = 30)%>%ts(start = c(2015,yday(g$Tarih[1])),
             frequency = 365),series = "30MA",size = 1.1)+
    autolayer(SMA(ts_object,n = 7)%>%ts(start = c(2015,yday(g$Tarih[1])),
             frequency = 365),series = "7MA",size = 1.1)+
    xlab("Date")+ylab("Demand(KwH)")+
    theme_economist()+
    theme(legend.position = "none")+
    ggtitle("Mean Daily Electric Demand from 2015 to 2022")
  
  p2 = ggAcf(ts_object,lag.max = 60,size = 1)+
    xlab("Lag")+ylab("Autocorrelation")+
    theme_fivethirtyeight()+
    ggtitle("Autocorrelation")
  
  p3 = ggPacf(ts_object,lag.max = 60,size = 1)+
    xlab("Lag")+ylab("Autocorrelation")+
    theme_fivethirtyeight()+
    ggtitle("Partial Autocorrelation")
  
  if(seasonality == "True")
  {
    return(grid.arrange(p1,s1,arrangeGrob(p2,p3,ncol = 2)))
  }else
  {
    return(grid.arrange(p1,arrangeGrob(p2,p3,ncol = 2)))
  }
    
  
}

plot_with(ts_object = ts_data,seasonality = "True")
```

When the seasonality graph is examined, there is seasonality in the data every year, but there is no obvious situation for the trend, we will look at it later.
In the pacf graph, on the other hand, the pacf value increases sharply at every 7th lag.

* Let's examine the trend variable with a regression model.

```{r}
tslm(ts_data~trend)%>%summary()
```

Since the trend variable is not significant, we can say that there is no obvious trend in the series.


## Check Stationary

```{r fig.height=10,fig.width=15,warning=FALSE,message=FALSE}
non_diff = ur.df(ts_data,selectlags = "AIC")%>%summary()
cat("Test Statistics for non Differencing : ",attr(non_diff,"teststat")[1],
       '\n',"Critical Values : ",attr(non_diff,"cval"))
one_diff = urca::ur.df(ts_data%>%diff(),selectlags = "AIC")%>%summary()
cat('\n',"Test Statistics for 1 Differencing : ",attr(one_diff,"teststat")[1],
       '\n',"Critical Values : ",attr(one_diff,"cval"))

plot_with(ts_object = diff(ts_data),seasonality = "False")
```

The series is not stationary. Therefore, we made it stationary with the 1st difference operation. This is a prerequisite for the Arima model.
According to the acf and pacf graphs of the difference series, I think the p (AR) parameter is 7, but I am not sure about the q (MA) parameter. I will try this.
Let's separate the last 2 weeks as the test set.


```{r}
train = window(ts_data,end = c(2022,122))
test = window(ts_data,start = c(2022,123))

model1 = Arima(train,order = c(7,1,6)) # I tried and found the most optimal q parameter to be 6.
predict1 = forecast(model1,
                   h = length(test),fan = T)

forecast::accuracy(predict1$mean,test) # mae 3329.389


model = auto.arima(train,
                   d = 1,
                   start.p = 5,
                   allowdrift = T,
                   allowmean = T,
                   seasonal = T,
                   stepwise = F,
                   approximation = F,
                   trace = T) # Auto arima 



predict = forecast(model,
                   h = length(test),fan = T)

forecast::accuracy(predict$mean,test)
```

```{r}
# Arima with seasonal and log transformation-Stepwise
model2 = auto.arima(log1p(train),
                   start.p = 5,
                   d = 1,
                   seasonal = T,
                   stepwise = T,
                   approximation = F,
                   trace = T,max.P = 7,start.P = 5)
predict2 = forecast(model2,
                   h = length(test),fan = F)

forecast::accuracy(predict2$mean%>%expm1(),test)
```

## Compare forecast

```{r,fig.height=6,fig.width=13}
autoplot(test,series = "Test",size = 1)+
  autolayer(predict1,series = "ARIMA(7,1,6)",PI = F,size = 1)+
  autolayer(predict$mean,series = "AutoArima(0,1,5)",PI = F,size = 1)+
  autolayer(predict2$mean%>%expm1(),series = "AutoArima(5,1,2)")+
  theme_hc()+
  ggtitle("Forecasting vs TestSeries")+ylab("Consumption Value")
```

If we look at the mae criterion and examine the graph, the ARIMA(7,1,6) model made the most compatible estimate.
No model has been able to capture the decline in the first 7 days well.

* Let's tabulate the results and look again

```{r}
acc_log_auto = forecast::accuracy(predict2$mean%>%expm1(),test)[3]
acc_auto = forecast::accuracy(predict$mean,test)[3]
acc_arima = forecast::accuracy(predict1$mean,test)[3]
result = data.frame("AutoWithLog(5,1,2)" = acc_log_auto,
                    "AutoArima(0,1,5)" = acc_auto,
                    "Arima(7,1,6)" = acc_arima,check.names = F)%>%t()%>%as.data.frame()

result
ggplot(result,aes(x = result%>%row.names(),y = V1,fill = result%>%row.names()))+
  geom_bar(stat = "identity")+
  xlab("Models")+ylab("MAE")+
  theme_fivethirtyeight()+
  theme(legend.position = "none")+
  ggtitle(label = "",subtitle = "Comparison of Arima Models(MAE)")+
  coord_flip()
  
```

We have determined the most compatible model.
Now let's model over the whole series and predict the next week.

```{r}
full_model = Arima(ts_data,
                   order = c(7,1,6))

predict = forecast(full_model,
                   h = 8,fan = F)

autoplot(window(ts_data,start = c(2022,1)),series = "History")+
  autolayer(predict$mean,series = "Forecast")+
  theme_hc()+
  ggtitle("Forecasting Next 7 Day")+ylab("Demand Value")
```


# Modeltime

Likewise, let's create estimates using the modeltime package.
I will split the data the same way and use the prophet, arima, arima boost and linear reg model here.

```{r,fig.height=8,fig.width=12}
g = g%>%as.data.frame()%>%select(Mean,Tarih)
training = g[(1:(dim(g)[1]-14)),]
testing = g[-(1:(dim(g)[1]-14)),]

arima_model = arima_reg()%>%
  set_engine("auto_arima")%>%
  fit(Mean~Tarih,training)

prophet_model = prophet_reg(seasonality_yearly = T,seasonality_daily = T)%>%
  set_engine("prophet")%>%
  fit(Mean~Tarih,data = training)

tslm_model = linear_reg()%>%
  set_engine("lm")%>%
  fit(Mean~as.numeric(Tarih)+factor(month(Tarih,label = T)),data = training)

arima_boost = arima_boost(learn_rate =.1 ,min_n = 5)%>%
  set_engine("auto_arima_xgboost")%>%
  fit(Mean~Tarih+factor(month(Tarih,label = T))+factor(day(Tarih)),data = training)
  

f_table = modeltime_table(
  arima_model,
  arima_boost,
  prophet_model,
  tslm_model
)

f_table%>%modeltime_calibrate(testing)%>%
  modeltime_accuracy()

```

According to the model results, the linear model gave the lowest mae value, but we can subtract it because it makes straight predictions.
The most compatible estimation model after the linear model was arima(2,0,0)(2,1,0).

and plotting

```{r}
f_table%>%modeltime_calibrate(testing)%>%
  modeltime_forecast(actual_data = testing)%>%
  plot_modeltime_forecast()
```


Except for the linear model, all models are close to each other and have the same pattern.
















